---
title: Effect of different water stress levels on morphophysiological variables of 15 crop and weed species.
author:
  - name: Chidimma Akabuogu 
  - name: Paul Bosse
  - name: Andrés Pascagaza
  - name: Javier Rojas
date: last-modified
date-format: iso
lang: en-EN
code-fold: true
toc: true
toc-location: body
toc-depth: 6
toc-title: Contents
format: html
editor: visual
---

# Team members

Chidimma Akabuogu, Paul Bosse, Andrés Pascagaza and Javier Rojas.

# Introduction

The global human population recently reached 8 billion people and the latest projections by the United Nations suggest that the world's population could grow to 8.5 billion by 2030 and 9.7 billion by 2050 (1). In a changing climate, environmental change is, therefore, an important factor for agriculture in an attempt to meet the growing need for food worldwide. Unfortunately, the vast majority of land is exposed to stressful conditions (2). Abiotic stress has been identified as one of the key factors threatening global food security. Compared to record yields, abiotic stress can reduce yields by more than 60% on average (3). Throughout their growth and development, crops encounter several abiotic factors or stresses (i.e. drought, flooding, osmotic stress, high temperatures, UV-radiation, nutrient imbalance, etc.) extremely affecting plant growth and development, leading to a reduction of the final yield and ultimately affecting worldwide food security (4).

Water stress has a direct impact on the development of living organisms of ecosystems. Plants rely on water for a proper growth and a lack of it can have a negative impact on agriculture. Drought is linked usually to a temporary lack of available water, and it usually affects arid and semiarid regions with lower precipitation, being some of them of relevance in terms of agricultural production. In case of a reduction in water availability the food productivity can be reduced significantly (5). Water stress can be caused by either natural or anthropogenic factors which include climate, soil salinity, agriculture or population demand. Soil salinity causes water stress in nearly 1 billion ha worldwide (6), by decreasing the osmotic pressure in the soil blocking water uptake by the rhizosphere (7).

As a result of water stress, plants exhibit several changes regarding morphology and physiology such as decline in height, alterations in the root to shoot ratio, reduction in nutrients uptake and their assimilation, among others. Some other effects related specifically to the photosynthesis process include the decrease in chlorophyll content and intercellular carbon dioxide. When it comes to the effects in the osmotic balance, plants go through a reduction of the turgor pressure and osmotic capacity caused by the accumulation of solutes due to water absence. Overall, and when all these effects are displayed together, the production and yield of the plant are drastically reduced (8).

However, the effects of water stress might variate according to the plants in concern. This project investigates the mechanisms of water stress tolerance in both crop and weed species. While water scarcity negatively affects weed growth, it does not entirely prevent seed germination, growth, and reproduction, resulting in a rise of weed seed production (9). Different cultivars respond differently to moisture stress, with some demonstrating resilience to extreme temperatures and unfavorable seed conditions. Moisture stress affects emergence, yield components and accelerates maturity, resulting in fewer tillers and seeds, ultimately reducing overall yield. The degree of these impacts is influenced by the growth stage, stress severity and cultivar genotype (10).

The aim of this project is to compare several morphophysiological variables of 15 different crop and weed species under three different water stress conditions. It is expected to contribute to the study of drought tolerance in order to optimize water management strategies and improve crop resilience to changing environmental conditions.

# Objectives and hypothesis

Objective: To compare the effect of different levels of water stress on physiological variables of a set of cultivated and wild plant species.

Hypothesis: The effect of water stress on the physiological variables of weed species will exhibit a better adaptation to water stress than the crop species.

# Methods

## Origin and structure of the data

Data comes from two different intake's experiments for the Adaptation to climate change course, in 2022 and 2023. The data was arranged by group, species, and week of the experiment. The variables were assessed depending on their treatment. The variables of interest for this study were: Plant height, soil humidity, leaf number, chlorophyll content, root length, aerial fresh weight, roots fresh weight, aerial dry weight, and roots dry weight. The data was retrieved from two collective databases by rendering the file called "00a_data-download.R". Then, the data from both analyzed years was bound by rendering the file called "00b_data-bind.R". Finally, the data was checked for any inconsistencies by rendering the file called "01_data-check.qmd". All the required files are found within this project file. The packages required for the data plotting and analysis included "ggplot2", "tidyr", "dplyr" and "agricolae".

## Graph plotting

Graphs were plotted displaying the mean values of the measured variables for both species and treatment. The variables measured along the whole time the experiment was conducted (soil humidity, plant height, and leaf number), line graphs were plotted in order to observe their progression. On the other hand, for variables measured just once at the end of the experiment (chlorophyll content, root length, aerial fresh weight, roots fresh weight, aerial dry weight, and roots dry weight), bar graphs were plotted.

## Statistical analysis

For each data set containing each one of the assessed variables for all the 15 species, a descriptive analysis was carried out by calculating the centered and deviation measures (mean, median, variance and standard error). In order to compare the different treatments, linear models were established for comparing more than two treatments. Both parametric (ANOVA and Tukey-HSD) and non-parametric (Kruskal-Wallis) tests were applied. The Shapiro-Wilk test for normality of the residuals was also applied, and according to the results was decided whether to use the results of parametric or non-parametric analysis.

# Results

```{r, include=FALSE}
# Load the data
load("data/bound-data.Rdata")
# Install libraries
# install and/or load the necessary libraries
# install.packages("ggplot2")
# install.packages("dplyr")
# install.packages("tidyr")
# install.packages("agricolae")
# install.packages("dplyr")
# install.packages("corrr")
# install.packages("ggcorrplot")
# install.packages("FactoMineR")
# install.packages("FactoMineR")
# install.packages("factoextra")
# install.packages("ggfortify")
# install.packages("kableExtra")
library(ggplot2)
library(dplyr)
library(tidyr)
library(agricolae)
library(dplyr)
library('corrr')
library(ggcorrplot)
library("FactoMineR")
library(factoextra)
library(ggfortify) 
library("kableExtra")

# Convert Week to a factor
datab$Week1 <- as.factor(datab$Week)
datab$Soil_humidity <-  as.numeric(datab$Soil_humidity)
datab$Species[datab$Species=="Amaranthus retroflexus "] <- "Amaranthus retroflexus" 
datab$Species[datab$Species=="Beta vulgaris "] <- "Beta vulgaris"
datab$Use[datab$Species=="Chenopodium album"] <- "wild"
datab$Use[datab$Species=="Amaranthus retroflexus"] <- "wild"
datab$Chlorophyll_content[datab$Chlorophyll_content=="211.4"] <- 21.4

#add variable of water content to the table
datab$Aerial_Water_Content <- (datab$Aerial_fresh_weight - datab$Aerial_dry_weight) 
datab$Root_Water_Content <- (datab$Roots_fresh_weight - datab$Roots_dry_weight)
```

```{r}
wild <- datab %>%
  filter(Use=="wild")
lev.wild <- levels(as.factor(wild$Species))
lev.wild

cult <- datab %>%
  filter(Use=="cultivated")
lev.cult <- levels(as.factor(cult$Species))
lev.cult
```

Plant height

check @fig-phw (naming chunk to check the figure in text)

```{r fig-phw, echo=TRUE, message=FALSE, warning=FALSE}
#| fig-cap: !expr paste("Mean", variablename ," of the wild species under three different water stress conditions.")
variable <- "Plant_height"
variablename <- "Plant height cm"
source("rmd/lineplotwild.R")
g
```

check @fig-phc (naming chunk to check the figure in text)

```{r fig-phc, echo=TRUE, message=FALSE, warning=FALSE}
#| fig-cap: !expr paste("Mean", variablename ," of the cultivated species under three different water stress conditions.")
variable <- "Plant_height"
variablename <- "Plant height cm"
source("rmd/lineplotcultivated.R")
g
```

Soil humidity

```{r fig-shw, echo=TRUE, message=FALSE, warning=FALSE}
#| fig-cap: !expr paste("Mean", variablename ," of the wild species under three different water stress conditions.")
variable <- "Soil_humidity"
variablename <- "Soil humidity"
source("rmd/lineplotwild.R")
g
```

```{r fig-shc, echo=TRUE, message=FALSE, warning=FALSE}
#| fig-cap: !expr paste("Mean", variablename ," of the cultivated species under three different water stress conditions.")
variable <- "Soil_humidity"
variablename <- "Soil humidity"
source("rmd/lineplotcultivated.R")
g
```

Leaf number

```{r fig-lnw, echo=TRUE, message=FALSE, warning=FALSE}
#| fig-cap: !expr paste("Mean", variablename ," of the wild species under three different water stress conditions.")
variable <- "Leaf_number"
variablename <- "Leaf number"
source("rmd/lineplotwild.R")
g
```

```{r fig-lnc, echo=TRUE, message=FALSE, warning=FALSE}
#| fig-cap: !expr paste("Mean", variablename ," of the cultivated species under three different water stress conditions.")
variable <- "Leaf_number"
variablename <- "Leaf number"
source("rmd/lineplotcultivated.R")
g
```

Chlorophyll content

```{r fig-ccw, echo=TRUE, message=FALSE, warning=FALSE}
#| fig-cap: !expr paste("Mean", variablesname, "of the wild species under three different water stress conditions.")
variables <- "Chlorophyll_content"
variablesname <- "Chlorophyll content (SPAD units)"
source("rmd/boxplot.R")
b
```

```{r fig-ccc, echo=TRUE, message=FALSE, warning=FALSE}
#| fig-cap: !expr paste("Mean", variablesname, "of the cultivated species under three different water stress conditions.")
variables <- "Chlorophyll_content"
variablesname <- "Chlorophyll content (SPAD units)"
source("rmd/boxplotcult.R")
b
```

Aerial fresh weight

```{r fig-afww, echo=TRUE, message=FALSE, warning=FALSE}
#| fig-cap: !expr paste("Mean", variablesname, "of the wild species under three different water stress conditions.")
variables <- "Aerial_fresh_weight"
variablesname <- "Aerial fresh weight (g)"
source("rmd/boxplot.R")
b
```

```{r fig-afwc, echo=TRUE, message=FALSE, warning=FALSE}
#| fig-cap: !expr paste("Mean", variablesname, "of the cultivated species under three different water stress conditions.")
variables <- "Aerial_fresh_weight"
variablesname <- "Aerial fresh weight (g)"
source("rmd/boxplotcult.R")
b
```

Aerial dry weight

```{r fig-adww, echo=TRUE, message=FALSE, warning=FALSE}
#| fig-cap: !expr paste("Mean", variablesname, "of the wild species under three different water stress conditions.")
variables <- "Aerial_dry_weight"
variablesname <- "Aerial dry weight (g)"
source("rmd/boxplot.R")
b
```

```{r fig-adwc, echo=TRUE, message=FALSE, warning=FALSE}
#| fig-cap: !expr paste("Mean", variablesname, "of the cultivated species under three different water stress conditions.")
variables <- "Aerial_dry_weight"
variablesname <- "Aerial dry weight (g)"
source("rmd/boxplotcult.R")
b
```

Root length

```{r fig-rlw, echo=TRUE, message=FALSE, warning=FALSE}
#| fig-cap: !expr paste("Mean", variablesname, "of the wild species under three different water stress conditions.")
variables <- "Root_length"
variablesname <- "Root length (cm)"
source("rmd/boxplot.R")
b
```

```{r fig-rlc, echo=TRUE, message=FALSE, warning=FALSE}
#| fig-cap: !expr paste("Mean", variablesname, "of the cultivated species under three different water stress conditions.")
variables <- "Root_length"
variablesname <- "Root length (cm)"
source("rmd/boxplotcult.R")
b
```

Roots fresh weight

```{r fig-rfww, echo=TRUE, message=FALSE, warning=FALSE}
#| fig-cap: !expr paste("Mean", variablesname, "of the wild species under three different water stress conditions.")
variables <- "Roots_fresh_weight"
variablesname <- "Roots fresh weight (g)"
source("rmd/boxplot.R")
b
```

```{r fig-rfwc, echo=TRUE, message=FALSE, warning=FALSE}
#| fig-cap: !expr paste("Mean", variablesname, "of the cultivated species under three different water stress conditions.")
variables <- "Roots_fresh_weight"
variablesname <- "Roots fresh weight (g)"
source("rmd/boxplotcult.R")
b
```

Roots dry weight

```{r fig-rdww, echo=TRUE, message=FALSE, warning=FALSE}
#| fig-cap: !expr paste("Mean", variablesname, "of the wild species under three different water stress conditions.")
variables <- "Roots_dry_weight"
variablesname <- "Roots dry weight (g)"
source("rmd/boxplot.R")
b
```

```{r fig-rdwc, echo=TRUE, message=FALSE, warning=FALSE}
#| fig-cap: !expr paste("Mean", variablesname, "of the cultivated species under three different water stress conditions.")
variables <- "Roots_dry_weight"
variablesname <- "Roots dry weight (g)"
source("rmd/boxplotcult.R")
b
```

Aerial water content

```{r fig-awcw, echo=TRUE, message=FALSE, warning=FALSE}
#| fig-cap: !expr paste("Mean", variablesname, "of the wild species under three different water stress conditions.")
variables <- "Aerial_Water_Content"
variablesname <- "Aerial Water Content (g)"
source("rmd/boxplot.R")
b
```

```{r fig-awcc, echo=TRUE, message=FALSE, warning=FALSE}
#| fig-cap: !expr paste("Mean", variablesname, "of the cultivated species under three different water stress conditions.")
variables <- "Aerial_Water_Content"
variablesname <- "Aerial Water Content (g)"
source("rmd/boxplotcult.R")
b
```

Root water content

```{r fig-rwcw, echo=TRUE, message=FALSE, warning=FALSE}
#| fig-cap: !expr paste("Mean", variablesname, "of the wild species under three different water stress conditions.")
variables <- "Root_Water_Content"
variablesname <- "Root Water Content (g)"
source("rmd/boxplot.R")
b
```

```{r fig-rwcc, echo=TRUE, message=FALSE, warning=FALSE}
#| fig-cap: !expr paste("Mean", variablesname, "of the cultivated species under three different water stress conditions.")
variables <- "Root_Water_Content"
variablesname <- "Root Water Content (g)"
source("rmd/boxplotcult.R")
b
```

## Statistical analysis - One-way ANOVA.

```{r, echo=TRUE, message=FALSE, warning=FALSE}
#| tbl-cap: !expr paste("Table 1. List of p-values obtained from one-way ANOVA for 15 species of cultivated and wild plants. Significant differences are highlighted in bold.")
p.value.ph <- numeric()
l <- levels(as.factor(datab$Species))
l <- l[-which(l=="Erigeron annuus")]
for(i in l) {
  d <- datab[datab$Species==i,]
  d <- d[!is.na(d$Week),]
  model.aov <- 
    aov(Plant_height ~ Treatment * Week +                        Error(PlantId/Week),data=d)
  s <- summary(model.aov)
  p.value.ph <- c(p.value.ph, 
                  s[[1]][[1]]$`Pr(>F)`[1])
}

p.value.ln <- numeric()
l <- levels(as.factor(datab$Species))
l <- l[-which(l=="Erigeron annuus")]
for(i in l) {
  d <- datab[datab$Species==i,]
  d <- d[!is.na(d$Week),]
  model.aov <- 
    aov(Leaf_number ~ Treatment * Week +                        Error(PlantId/(Treatment*Week)),data=d)
  s2 <- summary(model.aov)
  p.value.ln <- c(p.value.ln, 
                  s2[[1]][[1]]$`Pr(>F)`[1])
}

list.pvalues <- data.frame(l,p.value.ph,p.value.ln)
list.pvalues %>%
  kbl() %>%
  column_spec(2, bold = ifelse(list.pvalues$p.value.ph < 0.05, TRUE, FALSE)) %>%
  column_spec(3, bold = ifelse(list.pvalues$p.value.ln < 0.05, TRUE, FALSE)) %>% 
  kable_styling()
  

```

## Statistical analysis - Two-way ANOVA for repeated variables.

```{r, echo=TRUE, message=FALSE, warning=FALSE}
#| tbl-cap: !expr paste("Table 2. List of p-values obtained from two-way ANOVA for repeated variables for 15 species of cultivated and wild plants. Significant differences are highlighted in bold.")

# The following code was created for calculating the one way ANOVA for the variables that were measured just once. A table was created with all the p values for each species for each variable.

d <- datab %>%
  filter(!is.na(Chlorophyll_content),Week=="W6") %>%
  filter(!is.na(Aerial_fresh_weight),Week=="W6") %>%
  filter(!is.na(Aerial_dry_weight),Week=="W6") %>%
  filter(!is.na(Roots_fresh_weight),Week=="W6") %>%
  filter(!is.na(Roots_dry_weight),Week=="W6") %>%
  filter(!is.na(Root_length),Week=="W6")

p.value.cc <- numeric()
l <- levels(as.factor(d$Species))
for(i in l) {
  d1 <- d[d$Species==i,]
  model.aov <- 
    aov(formula=d1$Chlorophyll_content~d1$Treatment,data=d1)
  s3 <- summary(model.aov)
  p.value.cc <- c(p.value.cc, 
                  s3[[1]][["Pr(>F)"]][1])
}

p.value.afw <- numeric()
l <- levels(as.factor(d$Species))
for(i in l) {
  d1 <- d[d$Species==i,]
  model.aov <- 
    aov(formula=d1$Aerial_fresh_weight~d1$Treatment,data=d1)
  s4 <- summary(model.aov)
  p.value.afw <- c(p.value.afw, 
                  s4[[1]][["Pr(>F)"]][1])
}

p.value.adw <- numeric()
l <- levels(as.factor(d$Species))
for(i in l) {
  d1 <- d[d$Species==i,]
  model.aov <- 
    aov(formula=d1$Aerial_dry_weight~d1$Treatment,data=d1)
  s5 <- summary(model.aov)
  p.value.adw <- c(p.value.adw, 
                   s5[[1]][["Pr(>F)"]][1])
}

p.value.rfw <- numeric()
l <- levels(as.factor(d$Species))
for(i in l) {
  d1 <- d[d$Species==i,]
  model.aov <- 
    aov(formula=d1$Roots_fresh_weight~d1$Treatment,data=d1)
  s6 <- summary(model.aov)
  p.value.rfw <- c(p.value.rfw, 
                   s6[[1]][["Pr(>F)"]][1])
}

p.value.rdw <- numeric()
l <- levels(as.factor(d$Species))
for(i in l) {
  d1 <- d[d$Species==i,]
  model.aov <- 
    aov(formula=d1$Roots_dry_weight~d1$Treatment,data=d1)
  s7 <- summary(model.aov)
  p.value.rdw <- c(p.value.rdw, 
                   s7[[1]][["Pr(>F)"]][1])
}

p.value.rl <- numeric()
l <- levels(as.factor(d$Species))
for(i in l) {
  d1 <- d[d$Species==i,]
  model.aov <- 
    aov(formula=d1$Root_length~d1$Treatment,data=d1)
  s8 <- summary(model.aov)
  p.value.rl <- c(p.value.rl, 
                   s8[[1]][["Pr(>F)"]][1])
}

list.pvalues1 <- data.frame(l,p.value.cc,p.value.afw,p.value.adw,p.value.rfw,p.value.rdw,p.value.rl)

list.pvalues1 %>%
  kbl() %>%
  column_spec(2, bold = ifelse(list.pvalues1$p.value.cc < 0.05, TRUE, FALSE)) %>%
  column_spec(3, bold = ifelse(list.pvalues1$p.value.afw < 0.05, TRUE, FALSE)) %>%
  column_spec(4, bold = ifelse(list.pvalues1$p.value.adw < 0.05, TRUE, FALSE)) %>%
  column_spec(5, bold = ifelse(list.pvalues1$p.value.rfw < 0.05, TRUE, FALSE)) %>%
  column_spec(6, bold = ifelse(list.pvalues1$p.value.rdw < 0.05, TRUE, FALSE)) %>%
  column_spec(7, bold = ifelse(list.pvalues1$p.value.rl < 0.05, TRUE, FALSE)) %>%
  kable_styling()
```

### Statistical analysis - Principal Components Analysis (PCA).

```{r}
datab[ which(datab$Week=='W6'), ] #creates subset table with variables for week 6 to make PCA
#create two variables for the water content for aerial and root
pcadata <- datab[c(4:7,11,12,16,19,25,26)]

#show data
str(pcadata)
#check for empty values
colSums(is.na(pcadata))
#ATM only aerial water content and root water content are available for the PCA-------------DATA SHOULD BE FIXED!!
#fixing data
pcadata <- replace(pcadata, is.na(pcadata), 0) #change NA to 0
str(pcadata)
#Normalizing data
numerical_data <- pcadata[,5:10]
head(numerical_data)
data_normalized <- scale(numerical_data)#scale fuction applied
head(data_normalized)

corr_matrix <- cor(data_normalized)
ggcorrplot(corr_matrix,
           colors = c("pink", "purple", "darkred"))
#displays a correlation matrix
#data.pca <- princomp(corr_matrix)
data.pca <- princomp(data_normalized)
summary(data.pca)

data.pca$loadings[, 1:2]


fviz_cos2(data.pca, choice = "var", axes = 1:2)
pca2 <- fviz_pca_var(data.pca, col.var = "cos2",
            gradient.cols = c("darkred", "purple", "pink"),
            repel = TRUE)
pca2
#error The number of active individuals is different from the length of the factor habillage. Please, remove the supplementary individuals in the variable habillage.
pcafinal <- fviz_pca_biplot(data.pca,
                            label="var",
                            habillage = pcadata$Use,
                            addEllipses=TRUE,
                            ellipse.level=0.95,
                            repel = TRUE)
pcafinal
#pca dots IN PROGRESS NOT WORKING
# structure of the iris
# dataset
str(pcadata)

# print the data.pca dataset
head(pcadata)
data.pca <- prcomp(pcadata[,c(5:10)],
                   center = TRUE,
                   scale. = TRUE)

# summary of the
# prcomp object
summary(data.pca)

# structure of the pca object
str(data.pca)
#plotting pca

stress.pca.plot <- autoplot(data.pca,
                          data = pcadata,
                          colour = 'Species')
stress.pca.plot
```

# Discussion

The ANOVA results indicate that aerial fresh weight and aerial dry weight are the best variables for assessing water stress, based on their significant p-values. The data from the ANOVA does not provide evidence for comparing wild and cultivated plants. *Chenopodium album*, a wild species, demonstrated resistance in both conducted ANOVA tests. *Spinacia oleracea* and *Amaranthus retroflexus* may be tolerant as their ANOVA results show significant differences for only two of the assessed variables. *Raphanus sativus* might be a highly sensitive crop to drought stress, as it showed significant p-values for all assessed variables. Chlorophyll content may not be the most suitable variable for assessment, as it only showed significant values for four of the observed species. 

The PCA analysis revealed that root length was most strongly related to aerial and root water content. Additionally, there was a relation between aerial and root water content. While there is a slight relationship between root length and plant height, it is not as significant as the aforementioned variables. There is no significant relation between chlorophyll content and leaf number with the other variables.  Principal Component Analysis (PCA) related to cultivated and wild data showed that both overlapped, but the cultivated data had a stronger effect on the related variables mentioned before. On the other hand, the wild species had a stronger effect on chlorophyll and leaf number.

# Conclusions

The data analyzed in this study allows the identification of differences regarding the adaptive response of wild species in comparison to cultivated species. By exploring morphophysiological variables that are intrinsically linked to water stress tolerance, it was possible to identify that wild species exhibit an adaptive response towards an intermediate level of drought.

Among the assessed variables it has been found out that leaf number and plant height are the most suitable morphophysiological variables for the study of drought stress in the analyzed species, since with such variables it was possible to detect significant differences within the different groups of species and treatments. The species differ in their morphological values also due to their own water tolerance strategy. Further studies should be conducted in order to establish a window of tolerance to water stress, and including other important crop species.

# References

Boyer, J. S. (1982). Plant Productivity and Environment. In Science, 218, 443--448.

Cramer, G. R., Urano, K., Delrot, S., Pezzotti, M., & Shinozaki, K. (2011). Effects of Abiotic Stress on Plants: A Systems Biology Perspective. In BMC Plant Biology, 11, 1--14.

Donaldson, E. (1996). Crop traits for water stress tolerance. In American Journal of Alternative Agriculture, 11(2-3), 89-94. doi:10.1017/S0889189300006846.

Fahad, S., Bajwa, A. A., Nazir, U., Anjum, S. A., Farooq, A., Zohaib, A., et al. (2017). Crop production under drought and heat stress: Plant responses and management options. In Frontiers in Plant Science, 8, 29.

Ondrasek, G., Rengel, Z., & Veres, S. (2011). Soil Salinisation and Salt Stress in Crop Production. Abiotic Stress in Plants - Mechanisms and Adaptations. InTech. Available at: http://dx.doi.org/10.5772/22248.

Ondrasek, G. (2013). The Responses of Salt-Affected Plants to Cadmium. In P. Ahmad, M. M. Azooz, & M. N. V. Prasad (Eds.), Salt Stress in Plants. Springer, New York, NY. https://doi.org/10.1007/978-1-4614-6108-1_17.

Ondrasek, G. (2014). Water Scarcity and Water Stress in Agriculture. In P. Ahmad & M. Wani (Eds.), Physiological Mechanisms and Adaptation Strategies in Plants Under Changing Environment. Springer, New York, NY. <https://doi.org/10.1007/978-1-4614-8591-9_4>.

Singh, M., Thapa, R., Kukal, M. S., Irmak, S., Mirsky, S., & Jhala, A. J. (2022). Effect of Water Stress on Weed Germination, Growth Characteristics, and Seed Production: A Global Meta-Analysis. In Weed Science, 70(6), 621--640. DOI: 10.1017/wsc.2022.59.

Tajinder, K., & Saroj, A. (2021). Environmental Stress Physiology of Plants and Crop Productivity. Bentham Science Publishers Ltd.

United Nations. (2023). United Nations. \[Accessed on 29 November 2022\]. Available online: <https://www.un.org>.
